A Type System for String Sanitation\\Implemented Inside a Python

### 2014-04-16 sanitation-plas14/sanitation-plas14.tex
typo.
Fixed errors and typos in the rules.
Using new style for rules.

% Judgements
\newcommand{\err}{{\tt err}}
\newcommand{\trden}[1]{\llbracket #1 \rrbracket} % = Translation Denotation.

@@

%\authorinfo{~}{~}{~}
\authorinfo{Nathan Fulton\and Cyrus Omar\and Jonathan Aldrich}
           {School of Computer Science\\
            Carnegie Mellon University}
           {\{nfulton, comar, aldrich\}@cs.cmu.edu}

@@

This section presents the grammar and semantics of $\lcs$.
The semantics are defined in terms of an internal language with at least strings and a regex filter function.
These constraints are captured by the internal term valuations.
The internal language does not necessarily need a regex filter function because
any dynamic conversion is easily definable using a combination of filters and safe
casting.

% Whatever just dropping this in somewhere random.
\begin{thm}
If $S : \strin{r}$ then there exists a T such that $\trden{s} = T$ and either: 
\begin{itemize}
\item (a) $T \treduces \tstr{s}$ and $S \sreduces \strin{s}$, and $s \in lang{r}$.
\item (b) $T \err$ and $S \err$.
\end{itemize}
\end{thm}
\begin{proof}
By induction on the derivation of $\trden{S} = T$. TODO-nrf add case magic.
\begin{itemize}
\item Tr-string. By assumption $\trden{S} = \tstr{s}$. The (a) property follows.
By S-E-Str, $S \sreduces \str{s}$ and by T-E-Str, $\tstr{s} \treduces \tstr{s}$.
By source language type safety, it follows that $s \in \lang{r}$.
\item Tr-Concat.
By inversion of the typing relaiton for the source lnaguage, $S_1 : \strin{r_1}$ and $S_2 : \stringin{S_2}$.
By induction on the left premise, $T_1 \treduces \tstr{s_1}$, $S_1 \sreduces \str{s_1}$ and $s_1 \in \lang{r_1}$
or else $T_1 err$. 
Similarly for $T_2$ and $S_2$.
In the error cases, the propagations rules corresponding to S-E-Concat and T-E-Concat imply that $\tconcat{T_1}{T_2} \err$ and $\siconcat{S_1}{S_2} \err$.
In the non-error case, $\siconcat{S_1}{S_2} \sreduces \strin{s_1 \cdot s_2}$ by S-E-Concat and $\tconcat{T_1}{T_2} \treduces \tstr{s_1 \cdot s_2}$ by T-E-Concat.
Finally, $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ implies that $s_1 \cdot s_2 \in \lang{r}$ by \ref{lemmaregex}.
\item T-subst.
Suppose $\sisubst{r}{S_1}{S_2} : \strin{s}$ for some $x$. 
By inverison, $S_1 : \strin{r_1}$ and $S_2 : \strin{r_2}$.
Note that $s : \subst{r}{S_1}{S_2} : \stringin{\tsubst{r}{r_1}{r_2}}$ by S-T-Subst.
By induction, $\trden{S_1} = T_1$ and $\trden{S_2} = T_2$.
Either (a) or (b) holds for each.

%
% Begin Grammar 
%

@@

%The $\lcs$ language gives static semantics for common regular expression library
%functions. In this treatment, we include concatenation and filtering.
%The ${\tt filter}$ function removes all instances of a regular expression in a string,
%while concatenation ($+$) concatenates two strings.
%
%\subsection{Typing}
%
%The $\strin[r]$ type is parameterized by regular expressions; if $e:\strin[r]$,
%then $e \in r$. Mapping from an arbitrary $\str$ to a $\strin[r]$ requires
%defining an algorithm -- in terms of filter -- for converting a $\strin[.*]$
%into a $\strin[r]$. The static semantics of the language defines the types of
%operations on regular expressions in terms of well-understood properties about
%regular lanugages; we recall these properties in section 3.
%
%\subsection{Dynamics}
%
%There are two evaluation judgements: $e:T \reduces e'$ and $e:T \treduces i$.
%The $\reduces$ relation is between $\lcs$ expressions, while the $\treduces$
%relation is a mapping from $\lcs$ expressions into internal language expressions
%$i$ such that $i \ival$.
%
%Safety of the evaluation relation depends upon an injective mapping from $\lcs$ types info
%internal language types. This relation, $h$, is defined below.
%
%\subsection{Type Safety}
%
%The type safety proof relies upon some assumptions about the type system and
%dynamics of the internal language, as well as some properties of regular
%languages.
%
%There must exist a translation from $\lcs$ types to the types of the internal
%language. For the remainder of this paper, we call the type translation function
%$h$.
%
%\begin{defn}[Type Translation Function $h$]
%  The type translation function $h : Type \rightarrow IType$ is defined as follows:
%  \begin{itemize}
%    \item $\forall r. h(\strin[r]) = istr$
%    \item $h(\str) = istr$
%  \end{itemize}
%\end{defn}
%
%Additionally, we assume that the internal language contains an implementation of strings,
%together with operations for concatenation and filtering by regular expression.
%
%\begin{defn}[Types of internal values]
%  Let $\tstrf{s}$ range over string literals and $r$ over regular expressions.
%  Internal values are typed as follows:
%\begin{itemize}
%  \item If $e=\tstrf{s}$ then $e:istr$.
%  \item If $e=\ifilter{r}{\tstrf{s}}$ then $e:istr$.
%  \item If $e=\tstrf{s_1}+\tstrf{s_2}$ then $e:istr$.
%\end{itemize}
%\end{defn}
%
%For simplicity, we assume a fixed translation from
%$\lcs$ regular expressions to regular expressions recognizable by the internal
%language's regex library (in practice, a fixed translation is acceptable.)
%To summarize, we assume an internal language containing 
%a string type together with operations for string concatentation and filtering. We
%expect closure over strings for both operations.
%Finally, recall that ${\tt ifilter}$ is only needed for dynamic casts, which may be removed without
%descreasing the expressivity or even usability of the language.
%Finally, the semantics of the filter function are defined in terms of ${\tt rl\_filter}$,
%which is a static version of ${\tt ifilter}$.
%
\subsection{Properties of Regular Languages}

The regular languages are the smallest set generated by regular expressions
defined in Figure 1.

\begin{thm}{Closure Properties.} \label{thm:closure}
The regular expressions are closed under complements and concatenation.
\end{thm}
\begin{proof}
See \cite{cinderella}.
\end{proof}

\begin{thm}{Coercion Theorem.} \label{thm:coerce}
Suppose that $R$ and $L$ are regular expressions, and that $s \in R$ is a finite string.  Let $s' := coerce(R,L,s)$ with all maximal substrings recognized by $L$ replaced with $\epsilon$.  Then $s'$ is recognized by $(R \backslash L) + \emptyset$ and the construction of $R \backslash L$ is decidable.
\end{thm}
\begin{proof}
Let $F,G$ be FAs corresponding to $R$ and $L$, and let $G'$ be $G$ with its final states inverted (so that $G'$ is the complement of $L$).  Define an FA $H$ as a DFA corresponding to the NFA found by combining $F$ and $G'$ such that $H$ accepts only if $R$ and $L'$ accept or if $s$ is empty (this construction may result in an exponential blowup in state size.)  Clearly, $H$ corresponds to $R \backslash L + \emptyset$.  Thus, the construction of $R \backslash L + \emptyset$ is decidable.

@@

\onecolumn
\newcommand{\sctx}{\Gamma} % Context for external typing
\newcommand{\tctx}{\Omega} % Context for internal typing

@@

\begin{figure}
\begin{center}
\begin{tabular}{c r}

\inferline{S-T-include}
{\sctx \vdash \strin{s} : \stringin{r}}
{s \in \lang{r}}

\inferline{S-T-Concat}
{\sctx \vdash \siconcat{S_1}{S_2} : \strin{r_1 \cdot r_2}}
{\sctx \vdash S_1 : \strin{r_1} \\ \sctx \vdash S_2 : \strin{r_2}}

\inferline{S-T-subst}
{\sctx \vdash \sisubst{r}{S_1}{S_2} : r'}
{\sctx \vdash S_1 : \strin{r_1} \\ \sctx \vdash S_2 : \strin{r_2} \\ {\tt lsubst}(r,r_1,r_2) = r'}

\inferline{S-T-coerce}
{\sctx \vdash \coerce{r'}{s} : \stringin{r'}}
{\sctx \vdash S : \stringin{r}}

\end{tabular}
\caption{Typing rules for source langauge terms (S). The metavariable s ranges over
string literals.}
\label{fig:etyping}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{c r}
\inferline{S-E-Include}
{\strin{s} \sreduces \strin{s}}
{ \cdot }

\inferline{S-E-Concat}
{\siconcat{S_1}{S_2} \sreduces \strin{s_1 \cdot s_2}} % ???
{S_1 \sreduces \strin{s_1} \\ S_2 \sreduces \strin{s_2}}

\inferline{S-E-Subst}
{\sisubst{r}{S_1}{S_2} \sreduces s}
{S_1 \sreduces \strin{s_1} \\ S_2 \sreduces \strin{s_2} \\ \dosubst{r}{s_1}{s_2} = s}

\inferline{S-E-Coerce-OK}
{\coerce{r}{S} \sreduces \strin{s}}
{S \sreduces \strin{s} \\ x \in \lang{r}}

\inferline{S-E-Coerce-NotOk}
{S \err}
{\coerce{r}{S} \sreduces \strin{s} \\ s \not \in \lang{r}}

\end{tabular}
\caption{Big step semantics for the string fragment of the source language.
There are also unmentioned error propagation rules for for S-E-Concat and S-E-Subst.}
\label{fig:eeval}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{c r}

\inferline{P-T-String}
{\tctx \vdash \tstr{s} : \str}
{\cdot}

\inferline{P-T-Concat}
{\tctx \vdash \tconcat{P_1}{P_2} : \str}
{\tctx \vdash P_1 : \str \\ \tctx \vdash P_2 : \str}

\inferline{P-T-Subst}
{\tctx \vdash \tsubst{\rx{r}}{P_1}{P_2} : \str}
{\tctx \vdash P_1 : \str \\ \tctx \vdash P_2 : \str}

\inferline{P-T-Check}
{\tctx \vdash \tcheck{\rx{r}}{s} : \str}
{\tctx \vdash P : \str}

\end{tabular}
\caption{Typing rules for target langauge terms (P). The metavariable s ranges over
string literals.}
\label{fig:etyping}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{c r}
\inferline{P-E-Str}
{\tstr{s} \treduces \str{s}}
{ \cdot }

\inferline{P-E-Include}
{\rx{r} \treduces \rx{r}}
{ \cdot }

\inferline{P-E-Concat}
{\tconcat{P_1}{P_2} \treduces \tstr{s_1 \cdot s_2}} % ???
{P_1 \treduces \tstr{s_1} \\ P_2 \treduces \tstr{s_2}}

\inferline{P-E-Subst}
{\tsubst{r}{P_2}{P_3} \treduces \tstr{s}}
{P_1 \treduces \rx{r} \\ P_2 \treduces \tstr{s_2} \\ P_3 \treduces \tstr{s_3} \\ \dosubst{r}{s_2}{s_3} = s}

\inferline{P-E-Check-OK}
{\tcheck{r}{P} \treduces \tstr{s}}
{P \treduces \strin{s} \\ s \in \lang{r}}

\inferline{P-E-Check-NotOk}
{P \err}
{\tcheck{r}{T} \treduces \tstr{s} \\ s \not \in \lang{r}}

\end{tabular}
\caption{Big step semantics for the target lanugage.
There are also unimplemented error propagation rules for for Te-Concat and Te-Subst.}
\label{fig:ieval}
\end{center}
\end{figure}

%
% Begin big-step sematnics for internal terms.
%

\begin{figure}
\begin{center}
\begin{tabular}{c r}

\inferline{Tr-string}
{\trden{\strin{s}} = \tstr{s}}
{ \ }

\inferline{Tr-Concat}
{\trden{\siconcat{S_1}{S_2}} = \tconcat{P_1}{P_2}}
{ \trden{S_1} = P_1 \\ \trden{S_2} = P_2}

\inferline{Tr-Subst}
{\trden{ \sisubst{r}{S_1}{S_2} } = \tsubst{\rx{r}}{P_1}{P_2} }
{ \trden{S_1} = P_1 \\ \trden{S_2} = P_2 }

\inferline{Tr-Subst}
{\trden{ \coerce{r'}{S} } = P}
{ \trden{s} = P \\ S : \strin{r} \\ \lang{r'} \subseteq \lang{r}}

\inferline{Tr-Coerce-NotOk}
{\trden{\coerce{r'}{S}} = \tcheck{\rx{r'}}{S}}
{\trden{S} = P \\ s : \strin{r} \\ \lang{r'} \not \subseteq \lang{r}}

\end{tabular}
\caption{Translation from source terms (S) to target terms (P).
The translation is type-directed in the Tr-Coerce cases.}
\end{center}
\end{figure}

\twocolumn

@@

>>>>

% Judgements
\newcommand{\err}{\ {\tt err}}
\newcommand{\trden}[1]{\llbracket #1 \rrbracket} % = Translation Denotation.

@@

%\authorinfo{~}{~}{~}
\authorinfo{Nathan Fulton\and Cyrus Omar\and Jonathan Aldrich}
           {School of Computer Science\\
            Carnegie Mellon University}
           {\{nathanfu, comar, aldrich\}@cs.cmu.edu}

@@

This section presents the grammar and semantics of $\lcs$.
The semantics are defined in terms of an internal language with at least strings and a regex filter function.
These constraints are captured by the internal term valuations.
The internal language does not necessarily need a regex filter function because
any dynamic conversion is easily definable using a combination of filters and safe
casting.
%
% Begin Grammar 
%

@@

%The $\lcs$ language gives static semantics for common regular expression library
%functions. In this treatment, we include concatenation and filtering.
%The ${\tt filter}$ function removes all instances of a regular expression in a string,
%while concatenation ($+$) concatenates two strings.
%
%\subsection{Typing}
%
%The $\strin[r]$ type is parameterized by regular expressions; if $e:\strin[r]$,
%then $e \in r$. Mapping from an arbitrary $\str$ to a $\strin[r]$ requires
%defining an algorithm -- in terms of filter -- for converting a $\strin[.*]$
%into a $\strin[r]$. The static semantics of the language defines the types of
%operations on regular expressions in terms of well-understood properties about
%regular lanugages; we recall these properties in section 3.
%
%\subsection{Dynamics}
%
%There are two evaluation judgements: $e:T \reduces e'$ and $e:T \treduces i$.
%The $\reduces$ relation is between $\lcs$ expressions, while the $\treduces$
%relation is a mapping from $\lcs$ expressions into internal language expressions
%$i$ such that $i \ival$.
%
%Safety of the evaluation relation depends upon an injective mapping from $\lcs$ types info
%internal language types. This relation, $h$, is defined below.
%
%\subsection{Type Safety}
%
%The type safety proof relies upon some assumptions about the type system and
%dynamics of the internal language, as well as some properties of regular
%languages.
%
%There must exist a translation from $\lcs$ types to the types of the internal
%language. For the remainder of this paper, we call the type translation function
%$h$.
%
%\begin{defn}[Type Translation Function $h$]
%  The type translation function $h : Type \rightarrow IType$ is defined as follows:
%  \begin{itemize}
%    \item $\forall r. h(\strin[r]) = istr$
%    \item $h(\str) = istr$
%  \end{itemize}
%\end{defn}
%
%Additionally, we assume that the internal language contains an implementation of strings,
%together with operations for concatenation and filtering by regular expression.
%
%\begin{defn}[Types of internal values]
%  Let $\tstrf{s}$ range over string literals and $r$ over regular expressions.
%  Internal values are typed as follows:
%\begin{itemize}
%  \item If $e=\tstrf{s}$ then $e:istr$.
%  \item If $e=\ifilter{r}{\tstrf{s}}$ then $e:istr$.
%  \item If $e=\tstrf{s_1}+\tstrf{s_2}$ then $e:istr$.
%\end{itemize}
%\end{defn}
%
%For simplicity, we assume a fixed translation from
%$\lcs$ regular expressions to regular expressions recognizable by the internal
%language's regex library (in practice, a fixed translation is acceptable.)
%To summarize, we assume an internal language containing 
%a string type together with operations for string concatentation and filtering. We
%expect closure over strings for both operations.
%Finally, recall that ${\tt ifilter}$ is only needed for dynamic casts, which may be removed without
%descreasing the expressivity or even usability of the language.
%Finally, the semantics of the filter function are defined in terms of ${\tt rl\_filter}$,
%which is a static version of ${\tt ifilter}$.
%

@@

\newcommand{\sctx}{\Gamma} % Context for external typing
\newcommand{\tctx}{\Omega} % Context for internal typing

@@

\begin{figure}[t]
\small
$\fbox{\inferrule{}{\sctx \vdash S : \psi}}$
~~~~$\sctx ::= \emptyset \pipe \sctx, x : \psi$
\begin{mathpar}

\inferrule[S-T-include]
{s \in \lang{r}}
{\sctx \vdash \strin{s} : \stringin{r}}

\inferrule[S-T-Concat]
{\sctx \vdash S_1 : \stringin{r_1} \\ \sctx \vdash S_2 : \stringin{r_2}}
{\sctx \vdash \siconcat{S_1}{S_2} : \stringin{r_1 \cdot r_2}}

\inferrule[S-T-subst]
{\sctx \vdash S_1 : \stringin{r_1} \\ \sctx \vdash S_2 : \stringin{r_2} \\ {\tt lsubst}(r,r_1,r_2) = r'}
{\sctx \vdash \sisubst{r}{S_1}{S_2} : \stringin{r'}}

\inferrule[S-T-coerce]
{\sctx \vdash S : \stringin{r}}
{\sctx \vdash \coerce{r'}{s} : \stringin{r'}}

\end{mathpar}

\caption{Typing rules for source langauge terms (S). The metavariable s ranges over
string literals.}
\label{fig:etyping}
\end{figure}

\begin{figure}[t]
\small
$\fbox{\inferrule{}{S \sreduces S}}$
\begin{mathpar}

\inferrule[S-E-Include]
{ \ }
{\strin{s} \sreduces \strin{s}}

\inferrule[S-E-Concat]
{S_1 \sreduces \strin{s_1} \\ S_2 \sreduces \strin{s_2}} 
{\siconcat{S_1}{S_2} \sreduces \strin{s_1 \cdot s_2}} % ???

\inferrule[S-E-Subst]
{S_1 \sreduces \strin{s_1} \\ S_2 \sreduces \strin{s_2} \\ \dosubst{r}{s_1}{s_2} = s} 
{\sisubst{r}{S_1}{S_2} \sreduces s}

\inferrule[S-E-Coerce-OK]
{S \sreduces \strin{s} \\ s \in \lang{r}}
{\coerce{r}{S} \sreduces \strin{s}}

\inferrule[S-E-Coerce-NotOk]
{\coerce{r}{S} \sreduces \strin{s} \\ s \not \in \lang{r}}
{S \err}

\end{mathpar}
\caption{Big step semantics for the string fragment of the source language.
There are also unmentioned error propagation rules for for S-E-Concat and S-E-Subst.}
\label{fig:eeval}
\end{figure}

\newpage

\begin{figure}[t]
\small
$\fbox{\inferrule{}{\tctx \vdash P : \theta}}$
~~~~$\tctx ::= \emptyset \pipe \tctx, x : \theta$

\begin{mathpar}

\inferrule[P-T-String]
{ \ }
{\tctx \vdash \tstr{s} : \str}

\inferrule[P-T-Concat]
{\tctx \vdash P_1 : \str \\ \tctx \vdash P_2 : \str}
{\tctx \vdash \tconcat{P_1}{P_2} : \str}

\inferrule[P-T-Subst]
{\tctx \vdash P_1 : \str \\ \tctx \vdash P_2 : \str}
{\tctx \vdash \tsubst{\rx{r}}{P_1}{P_2} : \str}

\inferrule[P-T-Check]
{\tctx \vdash P : \str}
{\tctx \vdash \tcheck{\rx{r}}{s} : \str}

\end{mathpar}
\caption{Typing rules for target langauge terms (P). The metavariable s ranges over
string literals.}
\label{fig:etyping}
\end{figure}

\begin{figure}[t]
\small
$\fbox{\inferrule{}{P \treduces P}}$
\begin{mathpar}
\inferrule[P-E-Strin]
{ \ }
{\rx{r} \treduces \rx{r}}

\inferrule[P-E-Concat]
{P_1 \treduces \tstr{s_1} \\ P_2 \treduces \tstr{s_2}} 
{\tconcat{P_1}{P_2} \treduces \tstr{s_1 \cdot s_2}} % ???

\inferrule[P-E-Subst]
{P_1 \treduces \rx{r} \\ P_2 \treduces \tstr{s_2} \\ P_3 \treduces \tstr{s_3} \\ \dosubst{r}{s_2}{s_3} = s} 
{\tsubst{r}{P_2}{P_3} \treduces \tstr{s}}

\inferrule[P-E-Check-OK]
{P \treduces \strin{s} \\ s \in \lang{r}}
{\tcheck{r}{P} \treduces \tstr{s}}

\inferrule[P-E-Check-NotOk]
{\tcheck{r}{P} \treduces \tstr{s} \\ s \not \in \lang{r}}
{P \err}
\end{mathpar}
\caption{Big step semantics for the target lanugage.
There are also unimplemented error propagation rules for for P-E-Concat and P-E-Subst.}
\label{fig:ieval}
\end{figure}

\newpage

%
% Begin translation
%

\begin{figure}[t]

$\fbox{\inferrule{}{\trden{S} = P}}$
\begin{mathpar}

\inferrule[Tr-string]
{ \ }
{\trden{\strin{s}} = \tstr{s}}

\inferrule[Tr-Concat]
{ \trden{S_1} = P_1 \\ \trden{S_2} = P_2}
{\trden{\siconcat{S_1}{S_2}} = \tconcat{P_1}{P_2}}

\inferrule[Tr-Subst]
{ \trden{S_1} = P_1 \\ \trden{S_2} = P_2 }
{\trden{ \sisubst{r}{S_1}{S_2} } = \tsubst{\rx{r}}{P_1}{P_2} }

\inferrule[Tr-Coerce-Ok]
{ S : \strin{r} \\ \lang{r'} \subseteq \lang{r}}
{\trden{ \coerce{r'}{S} } = \tstr{s}}

\inferrule[Tr-Coerce-NotOk]
{\trden{S} = P \\ S : \strin{r} \\ \lang{r'} \not \subseteq \lang{r}}
{\trden{\coerce{r'}{S}} = \tcheck{\rx{r'}}{P}}

\end{mathpar}
\caption{Translation from source terms (S) to target terms (P).
The translation is type-directed in the Tr-Coerce cases.}
\end{figure}

@@

\subsection{Properties of Regular Languages}

The soundness theorems in this section depend upon some well-known properties
of regular expressions. Throughout this section, suppose regular expression
$r$ and strings $s$ are defined over a single, fixed, finite alphabet $\Sigma$.

\begin{thm} \label{thm:closure}
The regular expressions are closed under complements and concatenation.
\end{thm}
\begin{proof}
See \cite{cinderella}.
\end{proof}

\begin{defn}[$\tt{subst}$]
  The function $\subst{r}{s_1}{s_2}$ produces a string in which all substrings
  of $s_1$ matching $r$ are replaced with $s_2$.
\end{defn}

\begin{defn}[$\tt{lsubst}$]
  The function $\lsubst{r}{r_1}{r_2}$ produces a regular expression in which
  and sublanguage $r_1'$ of $r_1$ satisfying the condition $\lang{r_1'} \subseteq{r}$
  is replaced with $r_2$.
\end{defn}

We now present some properties about substitution.

\begin{thm}
  String substitution ($\subst{r}{s_1}{s_2}$) is decidable and defined for all expressions
  and strings over similar alpha.
\end{thm}

\begin{thm}
  Language substittion $(\lsubst{r}{r_1}{r_2})$ is decidable and defined for
  all expressions.
\end{thm}

The following relationship between string substituion and language substitution is at
the heart of our main correctness reesult.

\begin{thm}
  For all $r$, if $s_1 \in r_1$ and $s_2 \in r_2$ then 
  $\subst{r}{s_1}{s_2} \subseteq \lsubst{r}{r_1}{r_2}$.
\end{thm}

\begin{lem}{Coercion Lemma.} \label{thm:coercionalemma}
Suppose that $R$ and $L$ are regular expressions, and that $s \in R$ is a finite string.  Let $s' := coerce(R,L,s)$ with all maximal substrings recognized by $L$ replaced with $\epsilon$.  Then $s'$ is recognized by $(R \backslash L) + \emptyset$ and the construction of $R \backslash L$ is decidable.
\end{lem}
\begin{proof}
Let $F,G$ be FAs corresponding to $R$ and $L$, and let $G'$ be $G$ with its final states inverted (so that $G'$ is the complement of $L$).  Define an FA $H$ as a DFA corresponding to the NFA found by combining $F$ and $G'$ such that $H$ accepts only if $R$ and $L'$ accept or if $s$ is empty (this construction may result in an exponential blowup in state size.)  Clearly, $H$ corresponds to $R \backslash L + \emptyset$.  Thus, the construction of $R \backslash L + \emptyset$ is decidable.

\subsection{Safety of the Source and Target Languages}

We now prove safety for the string fragment of the source and target languages.

\begin{thm}
  Let $S$ be a term in the source language. If $\sctx \vdash S : \psi$ 
  then $S \sreduces S'$ and $S' : \psi$.
\end{thm}

\begin{thm}
  Let $P$ be a term in the target language. If $\sctx \vdash P : \theta$ 
  then $P \sreduces P'$ and $P' : \theta$.
\end{thm}

\subsection{Translation Correctness}

We now present the main correctness result.

\begin{thm}
If $S : \strin{r}$ then there exists a T such that $\trden{s} = T$ and either: 
\begin{itemize}
\item (a) $T \treduces \tstr{s}$ and $S \sreduces \strin{s}$, and $s \in lang{r}$.
\item (b) $T \err$ and $S \err$.
\end{itemize}
\end{thm}
\begin{proof}
By induction on the derivation of $\trden{S} = T$. TODO-nrf add case magic.
\begin{itemize}
\item Tr-string. By assumption $\trden{S} = \tstr{s}$. The (a) property follows.
By S-E-Str, $S \sreduces \str{s}$ and by T-E-Str, $\tstr{s} \treduces \tstr{s}$.
By source language type safety, it follows that $s \in \lang{r}$.
\item Tr-Concat.
By inversion of the typing relaiton for the source lnaguage, $S_1 : \strin{r_1}$ and $S_2 : \stringin{S_2}$.
By induction on the left premise, $T_1 \treduces \tstr{s_1}$, $S_1 \sreduces \str{s_1}$ and $s_1 \in \lang{r_1}$
or else $T_1 err$. 
Similarly for $T_2$ and $S_2$.
In the error cases, the propagations rules corresponding to S-E-Concat and T-E-Concat imply that $\tconcat{T_1}{T_2} \err$ and $\siconcat{S_1}{S_2} \err$.
In the non-error case, $\siconcat{S_1}{S_2} \sreduces \strin{s_1 \cdot s_2}$ by S-E-Concat and $\tconcat{T_1}{T_2} \treduces \tstr{s_1 \cdot s_2}$ by T-E-Concat.
Finally, $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ implies that $s_1 \cdot s_2 \in \lang{r}$ by \ref{lemmaregex}.
\item T-subst.
Suppose $\sisubst{r}{S_1}{S_2} : \strin{s}$ for some $x$. 
By inverison, $S_1 : \strin{r_1}$ and $S_2 : \strin{r_2}$.
Note that $s : \subst{r}{S_1}{S_2} : \stringin{\tsubst{r}{r_1}{r_2}}$ by S-T-Subst.
By induction, $\trden{S_1} = T_1$ and $\trden{S_2} = T_2$.
Either (a) or (b) holds for each.
